Here I attempt to replicate the results in @Cobb2012 using

```r
library(devtools) # devtools enables intsallation from alternate code locations
install_github("SODDr","noamross",ref=
```

```{r}
library(SODDr)
```


# Setup

First, I load a CSV file of parameters

```{r loadparms}
treeparms.df <- read.csv(system.file("paper_tree_parms_eq.csv",
                                     package="SODDr"), 
                         stringsAsFactors=FALSE)
print(treeparms.df)
```

This data set includes all of the model parameters as defined by @Cobb2012. 
Note that several values, such as the space occupied by different tanoak size
classes, and recruitment rates, are missing.  Per the original paper, I 
calculate these so as to maintain steady-state conditions in the model

```{r}
treeparms.df <- within(treeparms.df, {
# Calculate the relative space requirements of tanoak size classes
# based on initial conditions
space[1:4] <- 0.25*(S.init[1:4]/sum(S.init[1:4]))

# Set recruitment rates to steady-state levels.  For Redwood and Bay, this is 
# simply the mortality rate divided by the density-dependence coefficient at 
# simulation start.  For tanoak, which has multiple size classes, it's a but
# more involved

S.recruit[5] <- S.mortality[5]/(1-sum(S.init * space))
I.recruit[5] <- I.mortality[5]/(1-sum(S.init * space))
S.recruit[6] <- S.mortality[6]/(1-sum(S.init * space))
I.recruit[6] <- I.mortality[6]/(1-sum(S.init * space))

A2 <- S.transition[1]/(S.transition[2] + S.mortality[2])
A3 <- A2 * S.transition[2]/(S.transition[3] + S.mortality[3])
A4 <- A3* S.transition[3]/S.mortality[4]
S.recruit[1] <- (S.transition[1] + S.mortality[1])/(1-sum(S.init * space)) -
                (S.recruit[2] * A2 + S.recruit[3] * A3 + S.recruit[4] * A4)

I.recruit[1] <- S.recruit[1] 
})
```

Our parameters table defines the dispersal kernel as `adjacent.dispersal`, and
gives two parameters.  This function as outputs the first parameter
when within the cell, the second for dispersal to adjacent cells, and zero for
other cells:

```{r disperal kernel}
print(adjacent.dispersal)
```

Next I set up the locations in the model. These must be in the form of a matrix 
with the first column being site numbers, and the second being `x` and `y`
coordinates.  MakeLattice` is a convenient function that creates a regularly 
spaced grid:

```{r makelocations}
locations <- MakeLattice(nx=20,ny=20,dist=1)
head(locations)
```

Next, I create a matrix of initial population values, which should be sized by 
number of locations by number of species/size classes *times 2*.  They should be
in order of the classes as they appear in the data table, but alternating
$S,I,S,I\dots$. In this case, initial populations are uniform across the
landscape and equal to the `init` values from the parameters table.

```{r initial.values}
initial.vec = as.vector(rbind(treeparms.df$S.init, treeparms.df$I.init))
init <- matrix(data=initial.vec, nrow=nrow(locations), 
               ncol=2*nrow(treeparms.df), byrow=TRUE)
```

Finally, we set the time steps for the model

```{r}
time.steps <- 1:100
```

# Do it!

Running `SODModel` runs the model and generates a dataframe of population by
time, species, size class, and disease status

```{r}
pop.df <- SODModel(treeparms.df,locations,time.steps,init)
str(pop.df)
```


#Results
Here are the results, broken up by species and age class:

```{r}
library(ggplot2)
library(plyr)
pop.df.totals <- ddply(pop.df, c("Time", "Disease", "Species", "AgeClass"),
                       summarise, TotPop=sum(Population))
dynamic.plot <- ggplot(pop.df.totals, 
                       aes(x=Time,y=TotPop, fill=Disease, color=AgeClass)) + 
                  geom_area(position="stack", alpha=0.6) + facet_grid(~Species)
dynamic.plot
```

We can make a plot replicating the plot from @Cobb2012, showing only the
populations of small and large tanoaks as a proportion of the total population:

```{r}
paper.df <- ddply(transform(pop.df.totals, Size = ifelse(AgeClass == "1" | AgeClass == "2", "Small", "Large")), c("Time", "Species", "Size"), summarise, SizePop=sum(TotPop))
paper.df <- ddply(paper.df, "Time", transform, Pct=SizePop/sum(SizePop))
paper.plot <- ggplot(subset(paper.df, Species==1), aes(x=Time, y=Pct, lty=Size)) + geom_line(lwd=1) + scale_y_continuous(limits=c(0,0.6), expand=c(0,0)) + scale_linetype_manual(values=c(1,5))
paper.plot
```


